A User-Extensible and Adaptable Parser Architecture

### 2008-08-28 doc/sgai-2008.tex
Corrections from proofreading.

\abstract{%
    Some parsers need to be very precise and strict in what they parse, yet
    must allow users to easily adapt or extend the parser to parse new
    inputs, without requiring that the user have an in-depth knowledge and
    understanding of the parser's internal workings.  This paper presents a
    novel parsing architecture, designed for parsing Postfix log files,
    that aims to make the process of parsing new inputs as simple as
    possible, enabling users to trivially add new rules (to parse variants
    of existing inputs) and relatively easily add new actions (to process a
    previously unknown class of input).  The architecture scales linearly
    or better as the number of rules and/or size of input increases, making
    it suitable for parsing large corpora or months of accumulated data.
}

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental filters where
deficiencies are identified.  Most anti-spam techniques are
content-based~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms}
which requires accepting the mail in the first place to determine if it is
spam, but rejecting mail during the delivery attempt is preferable: senders
of non-spam mail that is mistakenly rejected will receive an immediate
non-delivery notice; resource usage is reduced on the accepting mail server
(allowing more intensive content-based techniques to be used); users have
less spam mail to wade through.  Improving the performance of anti-spam
techniques that are applied when mail is being transferred via
\SMTP{}\footnotemark{} is the goal of this project, by providing a platform
for reasoning about anti-spam filters.  The approach chosen to measure
performance is to analyse the log files produced by the \SMTP{} server in
use, Postfix~\cite{postfix}, rather than modifying it to generate
statistics: this approach improves the chances of other sites testing and
utilising the software.  The need arose for a parser capable of dealing
with the great number and variety of log lines produced by Postfix: the
parser must be designed so that adding support for parsing new inputs is a
simple task, as the log lines to be parsed will change over time.  The
variety in log lines occurs for several reasons:\footnotetext{Simple Mail
Transfer Protocol transfers mail across the Internet from the sender to one
or more recipients.  It is a simple, human readable, plain text protocol,
making it quite easy to test and debug problems with it.  The original
protocol definition is RFC~821~\cite{RFC821}, updated in
RFC~2821~\cite{RFC2821}.}

@@

    \item Log lines differ between versions of Postfix.

@@

It was hoped to reuse an existing parser rather than writing one from
scratch, but the existing parsers considered were rejected for one or more
reasons: they parsed too small a fraction of the log files; their parsing
was too inexact; they did not extract sufficient data.  The effort required
to adapt and improve an existing parser was judged to be greater than the
effort to write a new one, as the techniques used by the existing parsers
severely limited their potential: some ignored the majority of log lines,
parsing specific log lines with great accuracy, but without any provision
for readily parsing new or similar log lines; others sloppily parsed the
majority of log lines, but were incapable of distinguishing between log
lines of the same category, e.g.\ rejections.  The only prior published
work on the subject of parsing Postfix log files that the authors are aware
of is~\cite{log-mail-analyser}, which aims to present correlated log data
in a form suitable for a systems administrator to search using the myriad
of standard Unix text processing utilities already available.  A full state
of the art review is outside the scope of this paper but will be included
in the thesis resulting from this work; the relationship to another
procedural framework (ATN) is discussed below.

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of log lines (e.g.\ rejecting a mail), accompanied by a multitude
of precise patterns (\textit{rules\/}), each of which matches all log lines
of a specific type and only that type (e.g.\ rejection by a specific
\DNSBL{}).  It is an accepted standard to separate the parsing procedure
from the declarative grammar it operates with; part of the novelty here is
in the way that the grammar is itself partially procedural (each action is
a separate procedure).  This architecture is ideally suited to parsing
inputs where the input is not fully understood or does not conform to a
fixed grammar: the architecture warns about unparsed inputs and other
errors, but continues on parsing as best it can, allowing the developer of
a new parser to decide which deficiencies are most important and require
attention first, rather than being forced to fix the first error that
arises.

@@

Decoupling the parsing rules from their associated actions and the
framework allows new rules to be written and tested without requiring
modifications to the parser source code, significantly lowering the barrier
to entry for casual users who need to parse new inputs, e.g.\ part-time
systems administrators attempting to combat and reduce spam.  Decoupling
the actions from the framework simplifies both framework and actions: the
framework provides services to the actions, but does not need to perform
any tasks specific to the input being parsed; actions benefit from having
services provided by the framework, freeing them to concentrate on the task
of accurately and correctly processing the information provided by rules.

Decoupling also creates a clear separation of functionality: rules handle
low level details of identifying inputs and extracting data; actions handle
the higher level details of assembling the required data, dealing with the
intricacies of the input being parsed, complications arising, etc.;\ the
framework provides services to actions and manages the parsing process.

Some similarity exists between this architecture and William Wood's
\ATN{}~\cite{atns, nlpip}, a tool used in Computational Linguistics for
creating grammars to parse or generate sentences.  The resemblance between
the two (shown in \tableref{Similarities with ATN}) is accidental, but it
is obvious that the two approaches share a similar division of
responsibilities, despite having different semantics.

@@

\begin{table}[hbtp]
    \caption{Similarities with ATN}\label{Similarities with ATN}
    % The \smallskip{} below stop the text hitting the lines.
    \begin{tabular}[]{lll}
        \tabletopline{}%
        ATN           & Parser Architecture & Similarity                                \\
        \tablemiddleline{}%
        Networks      & Framework           & Determines the sequence of transitions or \\
                      &                     & actions that constitutes a valid input.   \\
        Transitions   & Actions             & Saves data and imposes conditions the     \\
                      &                     & input must meet to be accepted as valid.  \\
        Abbreviations & Rules               & Responsible for classifying input.        \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

Each action is a subroutine written to deal with a particular category of
input, e.g.\ rejections.  The actions are parser-specific: each parser
author will need to write the required actions from scratch unless
extending an existing parser.  It is anticipated that parsers utilising
this architecture will have a high ratio of rules to actions, with the aim
of having simpler rules and clearer distinctions between the inputs parsed
by different rules.  In the Postfix log parser developed for this project
there are \numberOFactions{} actions and \numberOFrules{} rules, with an
uneven distribution of rules to actions as shown in \graphref{Distribution
of rules per action}.
\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly, the action with
the most associated rules is \texttt{DELIVERY\_REJECTED}, the action that
handles Postfix rejecting a mail delivery attempt; it is followed by
\texttt{SAVE\_DATA}, the action responsible for handling informative log
lines, supplementing the data gathered from other log lines.  The third
most common action is, perhaps surprisingly, \texttt{UNINTERESTING}: this
action does nothing when executed, allowing uninteresting log lines to be
parsed without causing any effects (it does not imply that the input is
ungrammatical or unparsed).  Generally rules specifying the
\texttt{UNINTERESTING} action are rules to parse log lines that are not
associated with a specific mail, e.g.\ notices about configuration files
changing.  The remaining actions have only one or two associated rules:
some actions are required to address a deficiency in the log files, or a
complication that arises during parsing;  other actions will only ever have
one log line variant, e.g.\ all log lines showing that a remote client has
connected are matched by a single rule and handled by the \texttt{CONNECT}
action.

@@

The ability to add special purpose actions to deal with difficulties and
new requirements that are discovered during parser development is one of
the strengths of this architecture.  Instead of writing a single monolithic
function that must be modified to support new behaviour, with all the
attendant risks of adversely affecting the existing parser, when a new
requirement arises an independent action can be written to satisfy it.
Sometime the new action will require the cooperation of other actions,
e.g.\ to set or check a flag.  There is a possibility of introducing
failure when modifying existing actions in this way, but the modifications
will be smaller and occur less frequently than with a monolithic
architecture, thus failures will be less likely and will be easier to test
for and diagnose.  The architecture can be implemented in an object
oriented style, allowing sub-classes to extend or override actions in
addition to adding new actions; because each action is an independent
procedure, the sub-class need only modify the action it is overriding,
rather than reproducing large chunks of functionality.

During development of the Postfix log parser it became apparent that in
addition to the obvious variety in log lines, there were many complications
to be overcome.  Some were the result of deficiencies in Postfix's logging
(sometimes improved on or resolved by later versions of Postfix); others
were due to the vagaries of process scheduling, client behaviour, and
administrative actions.  All were successfully accommodated in the Postfix
log parser: adding new actions was enough to overcome several of the
complications; others required modifications to a single existing action to
work around the difficulties; the remainder were resolved by adapting
existing actions to cooperate and exchange extra data, changing their
behaviour as appropriate.

@@

Rules categorise inputs, specifying both the regex to match against each
input and the action to invoke when the match is successful.  The Postfix
log parser stores the rules it uses in the same SQL database the results
are stored in, removing any doubt about which set of rules was used to
produce a set of results; other implementations are free to store the rules
in whatever fashion suits their needs.  The framework warns about every
unparsed input, to alert the user that they need to alter or extend their
ruleset; the Postfix log parser successfully parses every log line in the
\numberOFlogFILESall{} log files it is currently being tested with.  The
framework requires each rule to have \texttt{action} and \texttt{regex}
attributes; each implementation is free to add any additional attributes it
requires.  The Postfix log parser adds several attributes: to optimise rule
ordering (\sectionref{Rule ordering}); to restrict which log lines each
rule can be matched against; and to describe each rule.  Data is
automatically extracted and saved from the log line based on keywords in
the rule's regex, but there is also a mechanism to specify extra data to be
saved.

Parsing new inputs is generally achieved by creating a new rule that pairs
an existing action with a new regex.  Adding new rules is simple: define
all the required attributes and append the new rule to wherever the
existing rules are stored.  The Postfix log parser supplies a utility based
on Simple Logfile Clustering Tool~\cite{slct-paper} to aid in producing
regexes from previously unparsed log lines.  Separating the rules from the
actions and framework enables other approaches to rule management to be
used, e.g.\ instead of manually adding new rules, machine learning
techniques could be used to automatically generate new rules based on the
existing rules and unparsed inputs.  If this approach was taken the choice
of machine learning technique would be constrained by the size of typical
data sets (see \sectionref{Results}).  Techniques requiring the full data
set when training would be impractical; Instance Based
Learning~\cite{instance-based-learning} techniques that automatically
determine which inputs from the training set are valuable and which inputs
can be discarded might reduce the data required to a manageable size.  This
avenue of research and development has not been pursued by the authors, but
could easily be undertaken independently.

The architecture does not try to detect overlapping rules: that
responsibility is left to the author of the rules.  Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the first matching rule wins, and the order in which rules are tried
against each input might change between parser invocations.  Overlapping
rules are frequently a requirement, allowing a more specific rule to match
some inputs and a more general rule to match the remainder, e.g.\
separating \SMTP{} delivery to specific sites from \SMTP{} delivery to the
rest of the world.  This can simplify both the general rule and the more
specific rule; additionally rules from different sources can be combined
with a minimum of prior cooperation or modification required.  Overlapping
rules should have a priority attribute to specify their relative ordering;
negative priorities may be useful for catchall rules.

Storing the rules separately to the parser allows external tools to be
written to detect overlapping rules.  Traditional regexes are equivalent in
computational power \FA{}, and can be converted to \FA{}, so regex overlap
can be detected by finding a non-empty intersection of two \FA{}\@.  The
standard equation for \FA{} intersection (given for example
in~\cite{intersection-of-NFA-using-Z}) is: $FA1 \cap{} FA2 =
\overline{(\overline{FA1} \cup{} \overline{FA2})}$.  Perl 5.10 regexes are
more powerful than traditional regexes, e.g.\ it is possible to match
correctly balanced brackets nested to an arbitrary depth:
\newline{}\tab{}\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!\newline{}
Perl 5.10 regexes can maintain an arbitrary state stack and are thus
equivalent in computational power to \PDA{} or \CFL{}, so detecting overlap
may require calculating the intersection of two \PDA{} or \CFL{}\@.  The
intersection of two \CFL{} is not closed, i.e.\ the resulting language
cannot always be parsed by a \CFL{}, so intersection may be
intractable in some cases e.g.:
$a^{*}b^{n}c^{n}~\cap~a^{n}b^{n}c^{*}~\rightarrow~a^{n}b^{n}c^{n}$.

Detecting overlap amongst $n$ regexes requires calculating $n(n-1)/2$
intersections, resulting in $O(n^2x)$ complexity, where $O(x)$ is the
complexity of calculating intersection.  This is certainly not a task to be
performed on every parsing run: detecting overlap amongst the Postfix log
parser's \numberOFrules{} rules would require calculating
\numberOFruleINTERSECTIONS{} intersections.

It is possible to define pathological regexes which fall into two main
categories: regexes that match every input, and regexes that consume
excessive amounts of CPU time during matching.  Defining a regex that
matches every input is trivial: \verb!/^/! matches the start of every
input.  Usually excessive CPU time is consumed when a regex with a lot of
alteration and variable quantifiers fails to match; successful matching is
generally quite fast.  This topic is beyond the scope of this paper,
see~\cite{mastering-regular-expressions} for in-depth discussion.

@@

    \item [Context-free rules:]  Rules do not take into account past or
        future inputs.  In context-free grammar terms the parser rules
        could be described as:
        \newline{}$\text{\textless{}input\textgreater{}} \mapsto
        \text{rule-1} | \text{rule-2} | \text{rule-3} | \dots |
        \text{rule-n}$

    \item [Partially context-sensitive actions and cascaded parsing:]
        Actions can consult the results (or lack of results) of previous
        actions during execution, providing some context sensitivity.
        Actions can not inspect past or future inputs, but actions can
        return a modified input to be parsed as if read from the input
        stream, allowing for a simplified version of cascaded
        parsing~\cite{cascaded-parsing}.

@@

    \item [Transduction:]  The parser can be thought of as implementing
        transduction: it takes data in one form (log files) and transforms
        it to another form (a database); other formats may be more suitable
        for other implementations.

@@

Parsing efficiency is an obvious concern when the Postfix log parser
routinely needs to parse large log files.  The mail server which generated
the log files used in testing the Postfix log parser accepts approximately
10,000 mails for 700 users per day; median log file size is 50MB,
containing 285,000 log lines --- large scale mail servers would have much
larger log files.  When generating the timing data used in this section,
\numberOFlogFILES{} log files (totaling 10.08 GB, \numberOFlogLINEShuman{}
log lines) were each parsed 10 times and the parsing times averaged.
Saving results to the database was disabled for the test runs, as the tests
are aimed at measuring the speed of the Postfix log parser rather than the
speed of the database.  The computer used for test runs is a Dell Optiplex
745 described in \tableref{Computer used to generate statistics}; it was
dedicated to the task of gathering statistics from test runs, and was not
used for any other purpose while test runs were ongoing --- any services
not necessary for running the tests were disabled.  Parsing all
\numberOFlogFILES{} log files in one run took
\input{build/include-timing-run-duration}, averaging
\input{build/include-timing-run-throughput.tex}.

@@

An important property of a parser is how parsing time scales relative to
input size: linearly, polynomially, or exponentially?  \Graphref{Parsing
time, log file size, and number of log lines} shows the parsing time in
seconds, log file size in MB, and number of log lines in tens of thousands,
for each of the \numberOFlogFILES{} log files.  All three lines on the
graph run roughly in parallel, giving the impression that the algorithm
scales linearly with input size.  This impression is borne out by
\graphref{parsing time vs log file size vs number lines factor}: the ratios
are tightly banded across the graph, showing that the algorithm scales
linearly.  The ratios increase (i.e.\ improve) for log files 22 and 62--68
despite their large size; that unusually large size is due to mail
forwarding loops resulting in a greatly increased number of mails delivered
and log lines generated.

@@

At the time of writing the Postfix log parser has \numberOFrules{}
different rules: as shown in \graphref{rule hits graph} 10\% of the rules
match the vast majority of log lines, with the remaining log lines split
across the other 90\% of the rules, similar to a Power Law distribution.
Assuming that the distribution of log lines is reasonably consistent over
time, parser efficiency should benefit from trying more frequently matching
rules before those which match less frequently.  To test this hypothesis
three full test runs were performed with different rule orderings:

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important for this architecture because it is
expected that typical parsers will have a large number of rules.  There are
\numberOFrules{} rules in the full Postfix log parser ruleset, whereas the
minimum number of rules required to parse the \numberOFlogFILES{} log files
is \numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the full
ruleset.  A second set of test runs was performed using the minimum
ruleset, and the parsing times compared to those generated using the full
ruleset: the percentage parsing time increase when using the full ruleset
instead of the minimal ruleset for optimal, shuffled and reverse orderings
is shown in \graphref{Percentage parsing time increase of maximum ruleset
over minimum ruleset}.  Clearly the increased number of rules has a
noticeable performance impact with reverse ordering, and a lesser impact
with shuffled ordering.  The optimal ordering shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset} in parsing time for a
\numberOFrulesMAXIMUMpercentage{} increase in the number of rules.  These
results show that the architecture scales very well as the number of rules
increases, and that optimally ordering the rules is an important
optimisation enabling this.
\showgraph{build/graph-full-ruleset-vs-minimum-ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}

>>>>

\abstract{%
    Some parsers need to be very precise and strict in what they parse, yet
    must allow users to easily adapt or extend the parser to parse new
    inputs, without requiring that the user have an in-depth knowledge and
    understanding of the parser's internal workings.  This paper presents a
    novel parsing architecture, designed for parsing Postfix log files,
    that aims to make the process of parsing new inputs as simple as
    possible, enabling users to trivially add new rules (to parse variants
    of existing inputs) and relatively easily add new actions (to process a
    previously unknown category of input).  The architecture scales
    linearly or better as the number of rules and size of input increases,
    making it suitable for parsing large corpora or months of accumulated
    data.
}

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental filters where
deficiencies are identified.  Most anti-spam techniques are content-based
(e.g.~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms}) requiring
that the mail be accepted before determining if it is spam, but rejecting
mail during the delivery attempt is preferable: senders of non-spam mail
that is mistakenly rejected will receive an immediate non-delivery notice;
resource usage is reduced on the accepting mail server (allowing more
intensive content-based techniques to be used on the mail that is
accepted); users have less spam mail to wade through.  Improving the
performance of anti-spam techniques that are applied when mail is being
transferred via \SMTP{}\footnotemark{} is the goal of this project, by
providing a platform for reasoning about anti-spam filters.  The approach
chosen to measure performance is to analyse the log files produced by the
\SMTP{} server in use, Postfix~\cite{postfix}, rather than modifying it to
generate statistics: this approach improves the chances of other sites
testing and utilising the software.  The need arose for a parser capable of
dealing with the great number and variety of log lines produced by Postfix:
the parser must be designed so that adding support for parsing new inputs
is a simple task, as the log lines to be parsed will change over time.  The
variety in log lines occurs for several reasons:\footnotetext{Simple Mail
Transfer Protocol transfers mail across the Internet from the sender to one
or more recipients.  It is a simple, human readable, plain text protocol,
making it quite easy to test and debug problems with it.  The original
protocol definition is RFC~821~\cite{RFC821}, updated in
RFC~2821~\cite{RFC2821}.}

@@

    \item Log lines differ amongst versions of Postfix.

@@

It was hoped to reuse an existing parser rather than writing one from
scratch, but the existing parsers considered were rejected for one or more
of the following reasons: they parsed too small a fraction of the log
files; their parsing was too inexact; they did not extract sufficient data.
The effort required to adapt and improve an existing parser was judged to
be greater than the effort to write a new one, because the techniques used
by the existing parsers severely limited their potential: some ignored the
majority of log lines, parsing specific log lines with great accuracy, but
without any provision for readily parsing new or similar log lines; others
sloppily parsed the majority of log lines, but were incapable of
distinguishing between log lines of the same category, e.g.\ rejections.
The only prior published work on the subject of parsing Postfix log files
that the authors are aware of is~\cite{log-mail-analyser}, which aims to
extract data from the log files, correlate it, and present it in a form
suitable for a systems administrator to search using the myriad of standard
Unix text processing utilities already available.  A full state of the art
review is outside the scope of this paper but will be included in the
thesis resulting from this work; the relationship to another procedural
framework (ATN) is discussed below.

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of inputs (e.g.\ rejecting a mail), accompanied by a multitude of
precise patterns (\textit{rules\/}), each of which matches all log lines of
a specific type and only that type (e.g.\ rejection by a specific
\DNSBL{}).  It is an accepted standard to separate the parsing procedure
from the declarative grammar it operates with; part of the novelty here is
in the way that the grammar is itself partially procedural (each action is
a separate procedure).  This architecture is ideally suited to parsing
inputs where the input is not fully understood or does not conform to a
fixed grammar: the architecture warns about unparsed inputs and other
errors, but continues on parsing as best it can, allowing the developer of
a new parser to decide which deficiencies are most important and require
attention first, rather than being forced to fix the first error that
arises.

@@

Decoupling the parsing rules from their associated actions allows new rules
to be written and tested without requiring modifications to the parser
source code, significantly lowering the barrier to entry for casual users
who need to parse new inputs, e.g.\ part-time systems administrators
attempting to combat and reduce spam; it also allows companies to develop
user-extensible parsers without divulging their source code.  Decoupling
the actions from the framework simplifies both framework and actions: the
framework provides services to the actions, but does not need to perform
any tasks specific to the input being parsed; actions benefit from having
services provided by the framework, freeing them to concentrate on the task
of accurately and correctly processing the information provided by rules.

Decoupling also creates a clear separation of functionality: rules handle
low level details of identifying inputs and extracting data; actions handle
the higher level tasks of assembling the required data, dealing with the
intricacies of the input being parsed, complications arising, etc.;\ the
framework provides services to actions and manages the parsing process.

Some similarity exists between this architecture and William Wood's
\ATN{}~\cite{atns, nlpip}, used in Computational Linguistics for creating
grammars to parse or generate sentences.  The resemblance between the two
(shown in \tableref{Similarities with ATN}) is accidental, but it is
obvious that the two approaches share a similar division of
responsibilities, despite having different semantics.

@@

\begin{table}[hbtp]
    \caption{Similarities with ATN}\label{Similarities with ATN}
    % The \smallskip{} below stop the text hitting the lines.
    \begin{tabular}[]{lll}
        \tabletopline{}%
        ATN           & Parser Architecture & Similarity                                \\
        \tablemiddleline{}%
        Networks      & Framework           & Determines the sequence of transitions or \\
                      &                     & actions that constitutes a valid input.   \\
        Transitions   & Actions             & Assembles data and imposes conditions the \\
                      &                     & input must meet to be accepted as valid.  \\
        Abbreviations & Rules               & Responsible for classifying input.        \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

Each action is a separate procedure written to deal with a particular category of
input, e.g.\ rejections.  The actions are parser-specific: each parser
author will need to write the required actions from scratch unless
extending an existing parser.  It is anticipated that parsers utilising
this architecture will have a high ratio of rules to actions, with the aim
of having simpler rules and clearer distinctions between the inputs parsed
by different rules.  In the Postfix log parser developed for this project
there are \numberOFactions{} actions and \numberOFrules{} rules, with an
uneven distribution of rules to actions as shown in \graphref{Distribution
of rules per action}.
\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly, the action with
the most associated rules is \texttt{DELIVERY\_REJECTED}, the action that
handles Postfix rejecting a mail delivery attempt; it is followed by
\texttt{SAVE\_DATA}, the action responsible for handling informative log
lines, supplementing the data gathered from other log lines.  The third
most common action is, perhaps surprisingly, \texttt{UNINTERESTING}: this
action does nothing when executed, allowing uninteresting log lines to be
parsed without causing any effects (it does not imply that the input is
ungrammatical or unparsed).  Generally rules specifying the
\texttt{UNINTERESTING} action are rules to parse log lines that are not
associated with a specific mail, e.g.\ notices about configuration files
changing.  The remaining actions have only one or two associated rules:
some actions are required to address a deficiency in the log files, or a
complication that arises during parsing;  other actions will only ever have
one log line variant, e.g.\ all log lines showing that a remote client has
connected are matched by a single rule and handled by the \texttt{CONNECT}
action.

@@

The ability to add special purpose actions to deal with difficulties and
new requirements that are discovered during parser development is one of
the strengths of this architecture.  Instead of writing a single monolithic
function that must be modified to support new behaviour, with all the
attendant risks of adversely affecting the existing parser, when a new
requirement arises an independent action can be written to satisfy it.
Sometimes the new action will require the cooperation of other actions,
e.g.\ to set or check a flag.  There is a possibility of introducing
failure when modifying existing actions in this way, but the modifications
will be smaller and occur less frequently than with a monolithic
architecture, thus failures will be less likely and will be easier to test
for and diagnose.  The architecture can be implemented in an object
oriented style, allowing sub-classes to extend or override actions in
addition to adding new actions; because each action is an independent
procedure, the sub-class need only modify the action it is overriding,
rather than reproducing large chunks of functionality.

During development of the Postfix log parser it became apparent that in
addition to the obvious variety in log lines there were many complications
to be overcome.  Some were the result of deficiencies in Postfix's logging
(some were improved on or resolved by later versions of Postfix); others
were due to the vagaries of process scheduling, client behaviour, and
administrative actions.  All were successfully accommodated in the Postfix
log parser: adding new actions was enough to overcome several of the
complications; others required modifications to a single existing action to
work around the difficulties; the remainder were resolved by adapting
existing actions to cooperate and exchange extra data, changing their
behaviour as appropriate.

@@

Rules categorise inputs, specifying both the regex to match against each
input and the action to invoke when the match is successful.  The Postfix
log parser stores the rules it uses in the same SQL database the results
are stored in, removing any doubt about which set of rules was used to
produce a set of results; other implementations are free to store their
rules in whatever fashion suits their needs.  The framework warns about
every unparsed input, to alert the user that they need to alter or extend
their ruleset; the Postfix log parser successfully parses every log line in
the \numberOFlogFILESall{} log files it is currently being tested with.
The framework requires each rule to have \texttt{action} and \texttt{regex}
attributes; each implementation is free to add any additional attributes it
requires.  The Postfix log parser adds several attributes: to optimise rule
ordering (\sectionref{Rule ordering}); to restrict which log lines each
rule can be matched against; and to describe each rule.  Data is
automatically extracted and saved from the input based on keywords in the
rule's regex, but the Postfix log parser provides a mechanism to specify
extra data to be saved.

Parsing new inputs is generally achieved by creating a new rule that pairs
an existing action with a new regex.  Adding new rules is simple: define
all the required attributes and append the new rule to wherever the
existing rules are stored.  The Postfix log parser supplies a utility based
on Simple Logfile Clustering Tool~\cite{slct-paper} to aid in producing
regexes from unparsed log lines.  Separating the rules from the
actions and framework enables other approaches to rule management to be
used, e.g.\ instead of manually adding new rules, machine learning
techniques could be used to automatically generate new rules based on the
existing rules and the unparsed inputs.  If this approach was taken the choice
of machine learning technique would be constrained by the size of typical
data sets (see \sectionref{Results}).  Techniques requiring the full data
set when training would be impractical; Instance Based
Learning~\cite{instance-based-learning} techniques that automatically
determine which inputs from the training set are valuable and which inputs
can be discarded might reduce the data required to a manageable size.  This
avenue of research and development has not been pursued by the authors, but
could easily be undertaken independently.

The architecture does not try to detect overlapping rules: that
responsibility is left to the author of the rules.  Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the first matching rule wins, and the order in which rules are tried
against each input might change between parser invocations.  Overlapping
rules are frequently a requirement, allowing a more specific rule to match
some inputs and a more general rule to match the remainder, e.g.\
separating \SMTP{} delivery to specific sites from \SMTP{} delivery to the
rest of the world.  Allowing overlapping rules simplifies both the general
rule and the more specific rule; additionally rules from different sources
can be combined with a minimum of prior cooperation or modification
required.  Overlapping rules should have a priority attribute to specify
their relative ordering; negative priorities may be useful for catchall
rules.

Decoupling the rules from the actions allows external tools to be written
to detect overlapping rules.  Traditional regexes are equivalent in
computational power to \FA{}, and can be converted to \FA{}, so regex
overlap can be detected by finding a non-empty intersection of two \FA{}\@.
The standard equation for \FA{} intersection (given for example
in~\cite{intersection-of-NFA-using-Z}) is: $FA1 \cap{} FA2 =
\overline{(\overline{FA1} \cup{} \overline{FA2})}$.  Perl 5.10 regexes are
more powerful than traditional regexes, e.g.\ it is possible to match
correctly balanced brackets nested to an arbitrary depth:
\newline{}\tab{}\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!\newline{}
Perl 5.10 regexes can maintain an arbitrary state stack and are thus
equivalent in computational power to \PDA{} or \CFL{}, so detecting overlap
may require calculating the intersection of two \PDA{} or \CFL{}\@.  The
intersection of two \CFL{} is not closed, i.e.\ the resulting language
cannot always be parsed by a \CFL{}, so intersection may be intractable in
some cases e.g.:
$a^{*}b^{n}c^{n}~\cap~a^{n}b^{n}c^{*}~\rightarrow~a^{n}b^{n}c^{n}$.

Detecting overlap amongst $n$ regexes requires calculating $n(n-1)/2$
intersections, resulting in $O(n^2x)$ complexity, where $O(x)$ is the
complexity of calculating intersection.  This is certainly not a task to be
performed every time the parser is used: detecting overlap amongst the Postfix log
parser's \numberOFrules{} rules would require calculating
\numberOFruleINTERSECTIONS{} intersections.

It is possible to define pathological regexes which fall into two main
categories: regexes that match every input, and regexes that consume
excessive amounts of CPU time during matching.  Defining a regex that
matches all inputs is trivial: \verb!/^/! matches the start of every input.
Usually excessive CPU time is consumed when a regex with a lot of
alteration and variable quantifiers fails to match; successful matching is
generally quite fast.  This topic is beyond the scope of this paper,
see~\cite{mastering-regular-expressions} for in-depth discussion.

@@

@@

    \item [Context-free rules:]  Rules can not take into account past or
        future inputs.  In context-free grammar terms the parser rules
        could be described as:
        \newline{}$\text{\textless{}input\textgreater{}} \mapsto
        \text{rule-1} | \text{rule-2} | \text{rule-3} | \dots |
        \text{rule-n}$

    \item [Context-aware actions:] Actions can consult the results (or lack
        of results) of previous actions during execution, providing some
        context sensitivity.

    \item [Cascaded parsing:] Actions can not inspect past or future
        inputs, but actions can return a modified input to be parsed as if
        read from the input stream, allowing for a simplified version of
        cascaded parsing~\cite{cascaded-parsing}.

    \item [Transduction:]  The architecture can be thought of as
        implementing transduction: it takes data in one form (log files)
        and transforms it to another form (a database); other formats may
        be more suitable for other implementations.

@@

Parsing efficiency is an obvious concern when the Postfix log parser
routinely needs to parse large log files.  The mail server which generated
the log files used in testing the Postfix log parser accepts approximately
10,000 mails for 700 users per day; median log file size is 50 MB,
containing 285,000 log lines --- large scale mail servers would have much
larger log files.  When generating the timing data used in this section,
\numberOFlogFILES{} log files (totaling 10.08 GB, \numberOFlogLINEShuman{}
log lines) were each parsed 10 times and the parsing times averaged.
Saving results to the database was disabled for the test runs, because the
tests are aimed at measuring the speed of the Postfix log parser rather
than the speed of the database.  The computer used for test runs is a Dell
Optiplex 745 described in \tableref{Computer used to generate statistics};
it was dedicated to the task of gathering statistics from test runs, and
was not used for any other purpose while test runs were ongoing --- any
services not necessary for running the tests were disabled.  Parsing all
\numberOFlogFILES{} log files in one run took
\input{build/include-timing-run-duration}, averaging
\input{build/include-timing-run-throughput.tex}.

@@

An important property of a parser is how parsing time scales relative to
input size: linearly, polynomially, or exponentially?  \Graphref{Parsing
time, log file size, and number of log lines} shows the parsing time in
seconds, log file size in MB, and number of log lines in tens of thousands,
for each of the \numberOFlogFILES{} log files.  The lines on the graph run
roughly in parallel, giving the impression that the algorithm scales
linearly with input size.  This impression is borne out by
\graphref{parsing time vs log file size vs number lines factor}: the ratios
are tightly banded across the graph, showing that the algorithm scales
linearly.  The ratios increase (i.e.\ improve) for log files 22 and 62--68
despite their large size; that unusually large size is due to mail
forwarding loops resulting in a greatly increased number of mails delivered
and log lines generated.

@@

At the time of writing the Postfix log parser has \numberOFrules{}
different rules: as shown in \graphref{rule hits graph} 10\% of the rules
XXX PERCENTAGE match the vast majority of log lines, with the remaining log lines split
across the other 90\% of the rules, similar to a Power Law distribution.
Assuming that the distribution of log lines is reasonably consistent over
time, parser efficiency should benefit from trying more frequently matching
rules before those which match less frequently.  To test this hypothesis
three full test runs were performed with different rule orderings:

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important for this architecture because it is
expected that typical parsers will have a large number of rules.  There are
\numberOFrules{} rules in the full Postfix log parser ruleset (it is tested
with \numberOFlogFILESall{} log files), whereas the
minimum number of rules required to parse the \numberOFlogFILES{} log files
is \numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the full
ruleset.  A second set of test runs was performed using the minimum
ruleset, and the parsing times compared to those generated using the full
ruleset: the percentage parsing time increase when using the full ruleset
instead of the minimal ruleset for optimal, shuffled and reverse orderings
is shown in \graphref{Percentage parsing time increase of maximum ruleset
over minimum ruleset}.  Clearly the increased number of rules has a
noticeable performance impact with reverse ordering, and a lesser impact
with shuffled ordering.  The optimal ordering shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset} in parsing time for a
\numberOFrulesMAXIMUMpercentage{} increase in the number of rules.  These
results show that the architecture scales very well as the number of rules
increases, and that optimally ordering the rules is an important
optimisation enabling this.
\showgraph{build/graph-full-ruleset-vs-minimum-ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}
